---
title: "Phackers Salary Survey"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

# Loading libraries and Data

```{r echo=FALSE}
library(tidyverse)
library(scales)  # allows us to use labels=comma in scales
```

```{r}
# read_csv is a tidyverse version, found in the readr library
phackers_raw <- read_csv("Philippine Tech Hackers Salary Survey 2017.csv")
dim(phackers_raw)
```
# Cleanup
We rename the columns to make them easier to work with. 

```{r}
phackers <- phackers_raw %>% rename(job_title=`What is your job title?`,
                    dept=`What is your department (if applicable)?`,
                    sal_raw=`Please enter your approximate monthly base salary in Philippine Pesos.`,
                    sal_increase_raw=`How much more is your current salary, than your salary three years ago?`,
                    sal_bonus_raw=`Approximately how much was your bonus from last year (aside from any mandatory bonuses -- e.g 13th month)?`,
                    in_ph=`Are you currently living/working in the Philippines?`,
                    ph_city=`If you currently work in the Philippines, which city do you work in?`,
                    age_range=`What is your age?`,
                    exp_raw=`How many years of experience do you have in your field?`,
                    gender=`What is your gender?`,
                    educ_stmts=`Which of the following statements about your education are true?`,
                    university=`Which university did you attend last?`,
                    university_other=`If your university was not listed in the previous question, you may add it here.`,
                    certifications=`What certifications do you currently have?`,
                    employment_status=`Are you a full- or part-time employee of a company, or a freelancer?`,
                    company_industry=`What is your company's business or industry?`,
                    company_size=`How many employees work at your company?`,
                    company_age=`How long has your company been in business?`,
                    os_used=`Operating systems`,
                    proglang_used=`Programming Languages`,
                    datatools_used=`Data tools`,
                    cloud_used=`Cloud/Containers`,
                    worktasks_noncollab=`Which of the following tasks play a part in your workday? [Writing code for non-collaborative projects (no one else will work on this code)]`,
                    worktasks_collab=`Which of the following tasks play a part in your workday? [Writing code for collaborative projects]`,
                    worktasks_collabreview = `Which of the following tasks play a part in your workday? [Reading/editing code originally written by others (e.g., using git)]`)
```
## Conversions
Convert string salary ranges to the numeric equivalent. 
```{r}
# phackers$sal_num <- parse_number(sapply(strsplit(phackers$sal_raw, " - "), function(x){x[1]}))
phackers$sal_num <- sapply(sapply(strsplit(phackers$sal_raw, ' - '), parse_number), mean)
phackers$exp_num <- parse_number(phackers$exp_raw)
```

Convert company size to factors so we can control order of facets
```{r}
# Convert company size to factors so we can control order of facets
# http://stackoverflow.com/questions/14262497/fixing-the-order-of-facets-in-ggplot
phackers$company_size_f <- factor(phackers$company_size, levels=c("1", "2 - 25", "25 - 100", "101 - 500", 
                                                                            "501- 1,000", "1,001 - 2,500", "2,501 - 10,000", "10,001 or more"))
# Convert gender to factor

phackers$company_industry_f <- factor(phackers$company_industry)
phackers$gender <- factor(phackers$gender, levels=c("Male", "Female", "Prefer not to say"))
# Convert company age to factor as well
# unique(phackers$company_age) will get us the list of ages to sort
phackers$company_age_f <- factor(phackers$company_age, levels=c("less than 2 years", "2 - 5 years",
                                                                          "6 - 10 years", "11 - 20 years", 
                                                                          "20 years or longer"))

# Convert the employment status to factors as well
phackers$employment_f <- factor(phackers$employment_status, levels=c("Employed in a company, full or part-time", "Freelance or independent contractor"), labels=c("Employed", "Freelance"))

# levels for age
l <- sort(unique(phackers$age_range))
print(l)
age_levels <- c(l[8], l[1:7])
print(age_levels)
phackers$age <- factor(phackers$age_range, labels = age_levels)

```

## Outlier labelling
We label as outliers those that:
 a. are outside the Philippines
 b. are outside Q3+3IQR in salary and years of experience
 
```{r}
sal_smry <- summary(phackers$sal_num)
sal_smry
```
```{r}
sal_iqr = sal_smry['3rd Qu.'] - sal_smry['1st Qu.']
sal_upper_fence <- sal_smry['1st Qu.'] + 3 * sal_iqr

exp_summary<-summary(phackers$exp_num)
exp_iqr <- exp_summary['3rd Qu.'] - exp_summary['1st Qu.']
exp_upper_fence <- exp_summary['3rd Qu.'] + 3*exp_iqr

phackers$outlier <- phackers$sal_num > sal_upper_fence | 
  phackers$exp_num > exp_upper_fence | 
  phackers$in_ph == FALSE

phackers <- phackers %>% filter(!outlier)
# From this point on, no need to filter(!outlier)
```
# Exploratory Data Analysis

## Gender, Age and Experience
```{r}
phackers %>% group_by(age, gender) %>% summarise(mean_sal = mean(sal_num), sd=sd(sal_num))
```
```{r  }
phackers %>% filter(gender != "Prefer not to say") %>% 
  ggplot(aes(x=exp_num, y=sal_num, color=gender)) + geom_jitter(alpha=0.5) +  
  geom_smooth(method="lm", se=FALSE) + scale_y_continuous(labels=comma) + xlab("Years of experience in field") +
  ylab("Salary")
```

```{r}
phackers %>% group_by(age) %>% 
  summarise(n=n(), "percent"=(round(100 * n()/sum(count(phackers)),2)))
```
```{r}
phackers %>% ggplot(aes(age, fill=gender)) + geom_bar(stat="count") + 
  ylab(label = "respondents") + ggtitle("Age Distribution", subtitle="by gender") 
```

```{r}
phackers %>% ggplot(aes(x=exp_num,fill=gender)) + geom_histogram(binwidth = 1) + 
  ylab(label="respondents") + xlab("years of experience")
```
## Where do they work? 
### Company Size and Age
```{r}
phackers %>% filter(employment_f=="Employed") %>% 
  group_by(company_size_f) %>% 
  summarize(n=n(), percentage=round(100*n()/sum(count(phackers)),2)) %>% 
  rename("Company Size"=company_size_f)
```
```{r}
phackers%>% filter(employment_f=="Employed") %>% ggplot(aes(company_size_f, fill=company_age_f)) + geom_bar(stat="count") +
  scale_fill_discrete("Company Age") + xlab("Company Size")
```
```{r}
phackers %>% filter(employment_f=="Employed") %>%  group_by(company_age_f) %>% summarize(n=n(), percentage=round(100*n()/sum(count(phackers)),2)) %>% rename("Company Age"=company_age_f)
```
```{r}
phackers %>% filter(employment_f=="Employed") %>% ggplot(aes(company_age_f, fill=company_size_f)) + geom_bar(stat="count") + 
  xlab("Company Age") + scale_fill_discrete("Company Size")
```
```{r}
unique(phackers$company_size)


phackers %>% filter(employment_f=="Employed") %>% ggplot(aes(y=sal_num, x=exp_num, color=company_size_f)) +
  geom_jitter(alpha=0.5) + facet_wrap(~company_age_f) + ylab("Salary") + xlab("Years of experience") +
  scale_color_discrete(name="Company Size") + labs(title="Employee Salaries", subtitle="by Company Size and Age")
  
```
## Industries
```{r}
phackers %>% filter(employment_f=="Employed") %>% group_by(company_industry_f) %>% 
  summarise(n=n(), pct=round(100*n()/sum(count(phackers)),2)) %>% arrange(desc(n)) %>%
  rename("Industry"=company_industry_f, percentage=pct)
```



## Gender distribution
```{r}
phackers %>% filter(!outlier, gender != "Prefer not to say") %>% ggplot(aes(x=gender, y=sal_num)) + 
#  geom_violin() + 
  geom_boxplot() +
  labs(title="Salary by gender", subtitle="Do women make less than men?", caption="This is a caption") +
  scale_y_continuous(labels =comma)

# phackers %>% count(gender, age)
```

```{r}
phackers %>% filter(gender!="Prefer not to say") %>% 
  select(sal_num, gender) %>% 
  ggplot(aes(y=sal_num, x=gender)) + 
  geom_boxplot() + 
  labs(title="Salary by gender", subtitle="Do women make less than men?", caption="This is a caption") + scale_y_continuous(labels =comma)
```


```{r out.width=9}
phackers %>% 
  filter(gender != "Prefer not to say") %>% 
  select(sal_num, exp_num, gender,employment_f ) %>% 
  ggplot(aes(exp_num, sal_num, color=gender, shape=employment_f),) + geom_jitter(alpha=0.5) + geom_smooth(method="lm", se=FALSE) + facet_grid(~employment_f)
```
```{r}
phackers %>% 
  select(sal_num, exp_num, gender) %>% 
  ggplot(aes(exp_num, sal_num)) + geom_jitter(alpha=.5) + geom_smooth(method = "lm")
```

```{r}
phackers %>% 
  select(sal_num, company_industry) %>% ggplot(aes(x=company_industry, y=sal_num)) + geom_boxplot() + geom_jitter()
  
```

# Test
```{r}
phackers %>% 
  filter(gender!="Prefer not to say") %>%
  select(sal_num, exp_num, gender,employment_status ) %>% 
  ggplot(aes(y=sal_num, x=gender),) + geom_boxplot() + facet_grid(~employment_status)
```
```{r}
phackers %>% filter(!outlier, employment_f=="Employed") %>% ggplot(aes(y=sal_num, x=company_size_f)) + geom_boxplot() + coord_flip() + xlab("Company Size") + ylab("Salary") + labs(title="Employees", subtitle="Full or part time")

```

```{r}
unique(phackers$company_age)

```

# Programming Languages
First, obtain the top 20 languages
```{r}
library(splitstackshape)
languages <- phackers %>% filter(!outlier) %>% select(proglang_used)
# cSplit_e(languages, "proglang_used", sep=";", fill=0, type="character")
language_counts <- cSplit(languages, "proglang_used", sep=";", direction="tall") %>% group_by(proglang_used) %>% summarise(cnt=n()) %>% arrange(desc(cnt)) %>% head(21) %>% filter(proglang_used != "(I don't use any programming language)")
language_counts
```
```{r}
top_20_languages <- factor(language_counts$proglang_used)
# Column names will be used later, for filtering
# we buld a list of column names. they all start with "proglang_used_" + language
top_20_colnames <- paste('proglang_used_',top_20_languages, sep="")
top_20_languages
```
## Linear model
```{r}
sal_langs <- phackers %>% filter(!outlier) %>% select(sal_num, proglang_used)

# Dummy variable for each of the top 20 programming languages
sal_lang_dummies <- concat.split.expanded(sal_langs, 2, sep=";", type="character", fill=0) %>% select(sal_num, one_of(top_20_colnames))
sal_lang_dummies
```

```{r}
lang_model <- lm(sal_num ~ ., sal_lang_dummies)
summary(lang_model)
```
```{r}
# we will cbind this later
phackers_languages <- sal_lang_dummies %>% select(-sal_num)
```

## University rankings
Top universities in SEAsia: https://www.topuniversities.com/university-rankings/asian-university-rankings/2016

Ranked universities: 
UP
ADMU
DLSU
UST
```{r}
# We create a new dummy variable (ranked_univ) that signifies whether it was one of those above
phackers <- phackers %>% mutate(univ_ranked = (university == "University of the Philippines Diliman, Quezon City" |
                                     university=="Ateneo de Manila University, Quezon City" | 
                                     university == "De La Salle University, Manila"|
                                     university == "University of Santo Tomas, Manila")) %>% filter(!outlier) 
phackers %>% select(sal_num, univ_ranked) %>%
  ggplot(aes(y=sal_num, x=univ_ranked)) + geom_boxplot() + scale_y_continuous(labels = comma)

```

## NCR vs. Outside NCR
```{r}
phackers <- phackers %>% mutate(in_ncr = endsWith(ph_city, " NCR")) 
phackers %>% filter(!outlier, !is.na(in_ncr)) %>% ggplot(aes(y=sal_num, x=in_ncr)) + geom_boxplot()
```

```{r}
phackers %>% mutate(in_ncr = endsWith(ph_city, " NCR")) %>% filter(!outlier & !is.na(in_ncr)) %>% ggplot(aes(y=sal_num, x=exp_num, color=in_ncr)) + geom_jitter() + scale_y_continuous(labels=comma) + geom_smooth(method="lm")
```

# Final Model
```{r}
phackers_vars <- cbind(phackers, phackers_languages) %>% filter(!outlier) %>% 
  select(sal_num, gender, exp_num, in_ncr, univ_ranked, company_age_f, company_size_f, 
         one_of(colnames(phackers_languages)) )
final_model <- lm(sal_num ~ . , data = phackers_vars)
summary(final_model)
```
```{r}
x_test <- phackers_vars[1:3,] %>% select(gender:proglang_used_Haskell)
predict(final_model, x_test)
```

```{r}
phackers_vars[1:3, "sal_num"]
```

# Where do they live and work
```{r}
phackers %>% filter(!is.na(ph_city) & ph_city!="I'm not currently working in the Philippines") %>% group_by(ph_city) %>% 
  summarise(c=n(), pct=round(100*n()/sum(count(phackers)),2), salary=mean(sal_num)) %>% 
  arrange(desc(c))
```
```{r}

phackers %>% filter(!is.na(ph_city) & ph_city!="I'm not currently working in the Philippines") %>% 
  reorder(ph_city)
  ggplot(aes(x=ph_city)) + geom_bar(stat="count") + coord_flip()
```

