---
title: "Philippine Tech Hackers Salary Survey, part 1"
output:
  html_document: 
    toc: true
    includes:
      in_header: meta_headers.html
    template: default_template.html
  
---


```{r intro, echo=FALSE, message=FALSE}
library(doMC)
registerDoMC(cores = 5)
library(tidyverse)
library(scales)  # allows us to use labels=comma in scales

source("salary_survey_util.R")

# read_csv is a tidyverse version, found in the readr library
phackers_raw <- read_csv("Philippine Tech Hackers Salary Survey 2017.csv")
phackers <- preprocess_salary_survey(phackers_raw)
#summary(phackers$salary_numeric)
#print(paste("Total non-outlier respondents: ", dim(phackers)[1]))

salary_median <- median(phackers$salary_numeric)
median_hline <-   geom_hline(yintercept=salary_median, linetype="dotted", color="blue")

```

# Introduction

The second Philippine Tech Hackers Salary Survey was conducted online using Google Forms. 
A total of `r nrow(phackers_raw)` community members responded, from companies both large and small, and from a wide variety of
industries. Respondents were mostly software developers, but other professionals who program and are involved in IT also participated. Outliers, and those who reported as living/working outside the country were excluded from the analysis.

The median monthly base salary for the remaining sample was `r Php(median(phackers$salary_numeric))`, with the middle half of all respondents earning between `r Php(quantile(phackers$salary_numeric, 0.25))` and `r Php(quantile(phackers$salary_numeric, 0.75))` 

_In the horizontal bar charts throughout this report, we include the interquartile range (IQR) to show the middle 50% of respondents’ answers to questions such as salary. One quarter of the respondents has a salary below the displayed range, and one quarter has a salary above the displayed range. The median salary of `r Php(median(phackers$salary_numeric))` will be represented in the box graphs by a dashed blue line._

```{r intro_plot, echo=FALSE, message=FALSE}
intro_plot <- ggplot(phackers, aes(salary_level, fill=..count..))  + 
  geom_bar() +  
  geom_vline(xintercept=47500, linetype="dashed", color="blue") + 
  labs(title="Total Salary", x="Total Salary (Php)", y="Respondents") +
  guides(fill = "none") +
  coord_flip() 


intro_plot
```

# Location

```{r location, echo=FALSE}
phackers_in_ph <- phackers %>%  filter(!is.na(ncr))
location_summary <-  phackers_in_ph %>% 
  group_by(ncr) %>% 
  summarise(Median=median(salary_numeric), n=n(), 
            pct=n()/sum(count(phackers_in_ph)), 
            ymin=quantile(salary_numeric, 0.25), 
            ymax=quantile(salary_numeric, 0.75))

median_salary_ncr <- location_summary$Median[location_summary$ncr == "NCR"]
percentage_outside_ncr <- location_summary$pct[location_summary$ncr != "NCR"]
median_salary_outside_ncr <- location_summary$Median[location_summary$ncr != "NCR"]
ncr_salary_premium <- median_salary_ncr - median_salary_outside_ncr
```

The overwhelming majority of respondents live and work in Metro Manila, with just over `r percent(percentage_outside_ncr)` of the respondents from other cities in the Philippines. 
The median salary for respondents living and working in Metro Manila is `r Php(median_salary_ncr)` with the median outside of NCR `r Php(ncr_salary_premium)` lower at `r Php(median_salary_outside_ncr)`

```{r location-plot, echo=FALSE}
location_lineplot <- location_summary %>% ggplot(aes(x=ncr, y=Median, ymin=ymin, ymax=ymax, fill=ncr)) + 
  median_hline + 
  geom_crossbar(alpha=0.5, color="white") +
   # geom_point(shape='|', size=10,stroke=2, color="white" )  +
  labs(title="Location", x="Location", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()

location_lineplot

# location_boxplot <- phackers_in_ph %>% ggplot(aes(x=ncr, y=salary_numeric, fill=ncr)) + 
#   median_hline + 
#   geom_boxplot() + 
#   labs(x="Location", y="Salary Range", title="Location") +
#   guides(fill="none") + 
#   coord_flip()
# location_plot
```

# Company Types 

```{r company_types, echo=FALSE}
phackers_employed <- phackers %>% filter(employment_f=="Employed")

# Determine the top 10 industries represented
top_industry_employed <- phackers_employed %>% 
  group_by(company_industry_f) %>% 
  summarise(median=median(salary_numeric), n=n(), pct=n()/sum(count(phackers))) %>% 
  arrange(desc(pct)) %>%
  rename("Industry"=company_industry_f, percentage=pct) %>% 
  top_n(n=10, wt=percentage)
```
The salary survey included questions on industry, company size, and company age. `r top_industry_employed$Industry[1]` was the most well-represented industry (`r percent(top_industry_employed$percentage[1])`), followed by `r top_industry_employed$Industry[2]`  (`r percent(top_industry_employed$percentage[2])`) and `r top_industry_employed$Industry[3]` (`r percent(top_industry_employed$percentage[3])`). 

## Industry Salaries
```{r industry_salaries, echo=FALSE}
top_industry_by_med_salary <- top_industry_employed %>% arrange(desc(median))
bottom_industry_by_med_salary <-top_industry_employed %>% arrange(median)
```

In terms of salary range, 
employees in `r top_industry_by_med_salary$Industry[1]` and `r top_industry_by_med_salary$Industry[2]` had the highest median salary at 
`r Php(top_industry_by_med_salary$median[1])` 
while  `r bottom_industry_by_med_salary$Industry[1]` had the lowest at `r Php(bottom_industry_by_med_salary$median[1])`.

```{r industry, echo=FALSE, message=FALSE}


# Use inner_join to get the subset phackers in the top 10 industries
# inner_join(phackers_employed, top_industry_employed, by=c("company_industry_f" = "Industry")) %>% 
#   group_by(company_industry_f) %>%
#   summarise(median(salary_numeric))

top_industry_phackers <- tbl_df(inner_join(phackers_employed, top_industry_employed, by=c("company_industry_f" = "Industry")) )

industry_boxplot <- top_industry_phackers %>% 
  ggplot(aes(x=company_industry_f, y=salary_numeric, fill=company_industry_f)) + 
  geom_boxplot() + 
  median_hline + 
  labs(title="Top 10 Industries by number of respondents", x="Industry", y="Salary Range") +
  guides(fill="none")  +
  coord_flip() 

top_industry_phackers %>% group_by(company_industry_f) %>% summarise(median_salary=median(salary_numeric), lower=quantile(salary_numeric, 0.25), upper=quantile(salary_numeric, 0.75)) %>%
  ggplot(aes(x=company_industry_f, fill=company_industry_f, y=median_salary, ymin=lower, ymax=upper)) + 
  median_hline + 
  # geom_linerange(size=5) +
  geom_crossbar(alpha=0.5, color="white") + 
  # geom_point(shape='|', size=5, stroke=2, color="white" )  +
  labs(title="Company Industry", x="Company Industry", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```

## Company Size 


```{r company_size, echo=FALSE}

company_size_summary <- phackers_employed %>%
  group_by(company_size_f) %>% 
  summarise(median_salary=median(salary_numeric), n = n(), pct = n() / sum(count(phackers)),
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, 0.75))
```

Half the respondents work in companies with 100 or less employees, 
with `r percent(company_size_summary$pct[company_size_summary$company_size_f=="2 - 25"])` of respondents working in companies with between 2 and 25 employees. 
Respondents working in very large companies (over 10,000 employees) made up only `r percent(company_size_summary$pct[company_size_summary$company_size_f=="10,001 or more"])` of the sample. 

```{r company_size_plot, echo=FALSE}
company_size_summary %>% ggplot(aes(x=company_size_f, y=median_salary, ymin=lower, ymax=upper, fill=company_size_f )) + 
  median_hline + 
  geom_crossbar(alpha=0.5, color="white") + 
  # geom_linerange(size=5) +
  # geom_point(shape='|', size=5, stroke=2, color="white" )  +
  labs(title="Company Size", x="Company Size", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```

## Company Age


```{r company_age, echo=FALSE}
company_age_summary <- phackers_employed %>%
  group_by(company_age_f) %>% 
  summarise(median_salary=median(salary_numeric), n(), 
            pct = n()/sum(count(phackers)),
            lower=quantile(salary_numeric, 0.25),
            upper=quantile(salary_numeric, 0.75))
```

Employees working at companies less than 10 years old made up 
`r percent(sum(company_age_summary$pct[company_age_summary$company_age_f <= "6 - 10 years"]))` of respondents 
and reported salaries above the overall median.

```{r company_age_plot, echo=FALSE}
company_age_boxplot <- phackers_employed %>% 
  ggplot(aes(x=company_age_f, y=salary_numeric, fill=company_age_f)) + 
  median_hline +
  geom_boxplot(outlier.alpha="0") +
  labs(x="Company age", y="Salary Range", title="Company Age") +
  guides(fill="none") +
  coord_flip() 

company_age_summary %>% ggplot(aes(x=company_age_f, fill=company_age_f, y=median_salary, ymin=lower, ymax=upper)) + 
  median_hline +
  geom_crossbar(alpha=0.5, color="white") +
  labs(title="Company Age", x="Company Age", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```

# Team and Project Size
There seems to be no clear relationship between team or project size and salary, other than solo developers and those on
very large project teams (greater than 20 members) tend to make less than the overall median salary.

```{r team_project_size, echo=FALSE}
team_size_summary <- phackers_employed %>% 
  group_by(team_size_f) %>% 
  summarise(y=median(salary_numeric), n=n(), percent=round(100*n()/sum(count(phackers)),2),
            lower=quantile(salary_numeric, 0.25),
            upper=quantile(salary_numeric, 0.75))

team_size_boxplot <- phackers_employed %>% 
  ggplot(aes(x=team_size_f, y=salary_numeric, fill=team_size_f)) +
  median_hline + 
  geom_boxplot() + 
  guides(fill="none") + 
  labs(title="Team Size", x="Team size", y="Salary Range")

team_size_summary %>% ggplot(aes(x=team_size_f, fill=team_size_f, y=y, ymin=lower, ymax=upper)) +
  median_hline +
  geom_crossbar(alpha=0.5, color="white") + 
  # geom_linerange(size=5) +
  # geom_point(shape='|', size=5, stroke=2, color="white" )  +
  labs(title="Team Size", x="Team Size", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
  
project_size_summary <- phackers_employed %>% group_by(project_size_f) %>% 
  summarise(y=median(salary_numeric), 
            n=n(), 
            percent=round(100*n()/sum(count(phackers)),2),
            lower=quantile(salary_numeric, 0.25),
            upper=quantile(salary_numeric, 0.75))

project_size_boxplot <- phackers_employed %>%
  ggplot(aes(x=project_size_f, y=salary_numeric, fill=project_size_f)) + 
  median_hline + 
  geom_boxplot() + 
  guides(fill="none") +
  labs(title="Project Size", x="Project Size", y="Salary Range")

project_size_summary %>% ggplot(aes(x=project_size_f, fill=project_size_f, y=y, ymin=lower, ymax=upper)) +
  median_hline +
  # geom_linerange(size=5) +
  geom_crossbar(alpha=0.5, color="white") + 
  labs(title="Project Size", x="Project Size", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```

# Freelance and Independent Contractors


```{r employment, echo=FALSE}
freelance_salary_summary <- phackers %>% group_by(employment_f) %>% 
  summarise(median_salary=median(salary_numeric), 
            n=n(), 
            percent=n()/sum(count(phackers)),
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, 0.75))

freelance_premium <- freelance_salary_summary$median_salary[freelance_salary_summary$employment_f == "Freelance"] -
  freelance_salary_summary$median_salary[freelance_salary_summary$employment_f == "Employed"]
```

Freelancers or independent contractors made up only `r percent(freelance_salary_summary$percent[freelance_salary_summary$employment_f == "Freelance"])` of the respondents. As a group, freelancers enjoyed a `r Php(freelance_premium)` premium over employees.

```{r employment_plot, echo=FALSE}
freelance_salary_summary %>% ggplot(aes(x=employment_f, fill=employment_f, y=median_salary, ymin=lower, ymax=upper)) +
  median_hline +
  geom_crossbar(alpha=0.5, color="white") +
  labs(title="Employment", x="Employment", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```


```{r, echo=FALSE}
phackers_freelance = phackers %>% filter(employment_f == "Freelance")

phackers_freelance_summary <- phackers_freelance %>%  group_by(company_industry_f) %>% 
  summarise(median=median(salary_numeric), n=n(), pct=round(100*n()/sum(count(phackers_freelance)),2)) %>% 
  arrange(desc(pct)) %>%
  rename("Industry"=company_industry_f, percentage=pct) %>% 
  top_n(n=10, wt=percentage)

# Not porinting this because almost all of freelancers work in Software, and their median is same as the overall median. Only 3 work in consulting, 2 in BPO and 
# other industries only have 1 Freelance respondent
# phackers_freelance_summary
```

```{r, echo=FALSE, message=FALSE}
d <- phackers %>% select(salary_numeric, ncr, employment_f, company_age, company_size_f, company_industry_f, age_range_f, exp_num, university)
model = lm(salary_numeric ~ ., data=d)
```