---
title: "Philippine Tech Hackers Salary Survey, part 2"
output:
  html_document: 
    toc: true
    includes:
      in_header: meta_headers2.html
    template: default_template.html
  html_notebook: default
  pdf_document: default
---

```{r echo=FALSE, message=FALSE}
library(doMC)
registerDoMC(cores = 5)

library(tidyverse)
library(scales)  # allows us to use labels=comma in scales

source("salary_survey_util.R")

# read_csv is a tidyverse version, found in the readr library
phackers_raw <- read_csv("Philippine Tech Hackers Salary Survey 2017.csv")
phackers <- preprocess_salary_survey(phackers_raw)

salary_median <- median(phackers$salary_numeric)
median_hline <-   geom_hline(yintercept=salary_median, linetype="dotted", color="blue")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Gender
```{r gender, echo=FALSE, echo=FALSE}
gender_summary <- phackers %>% group_by(gender) %>% summarise(n=n(), 
                                                              percent=n()/sum(count(phackers)),
                                                              median=median(salary_numeric), 
                                                              lower=quantile(salary_numeric, 0.25), 
                                                              upper=quantile(salary_numeric, 0.75))
```

The sample was overwhelmingly male (`r percent(gender_summary$percent[gender_summary$gender=="Male"])`), and while the median salary of both male and female respondents were the same, the upper bound of the middle 50% of salaries for men exceeded women's by `r Php(gender_summary$upper[gender_summary$gender=="Male"] - gender_summary$upper[gender_summary$gender=="Female"])`.

```{r gender_plot, echo=FALSE, message=FALSE}
# Plot
gender_summary %>% filter(gender != "Prefer not to say") %>% 
  ggplot(aes(x=gender, y=median, ymin=lower, ymax=upper, fill=gender)) +  median_hline +
  geom_crossbar(alpha=0.5, color="white") + 
  geom_jitter(alpha=I(0.1)) + 
  guides(fill="none") + 
  labs(x="Gender", y="Salary Range", title="Gender") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  coord_flip()
```


## Age and Experience

```{r age, echo=FALSE, message=FALSE}
age_hist <- phackers %>% 
  ggplot(aes(x=age_range_f, fill=..count..)) + 
  geom_bar(stat="count") + 
  labs(title="Age", x="Age Group", y="")+
  guides(fill="none")

age_hist

phackers_by_age <- phackers %>% group_by(age_range_f) %>% summarise(n=n(), percent=n()/sum(count(phackers)))
phackers_under_25 <- sum(phackers_by_age$percent[phackers_by_age$age_range_f <= "21-25"])
```

Over half (`r percent(phackers_under_25)`) of respondents were 25 years old or younger at the time of the survey. There were not enough respondents with ages greater than 40 to provide meaningful analysis. As expected, developer salary increases with age and experience. 

```{r age_under_40, echo=FALSE, message=FALSE}
# Not enough in the ranges > 40yo to plot meaningful stats, so we restrict analysis to the ranges 40 and below
phackers_under_40 <- phackers %>% filter(age_range_f < "41-45") 
summary_age <- phackers_under_40 %>%  group_by(age_range_f) %>% 
  summarise(n=n(), 
            median_salary=median(salary_numeric), 
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, .75)) 

# Graph using geom_linerange
summary_age %>% ggplot(aes(x=age_range_f, y=median_salary, fill=age_range_f, ymin=lower, ymax=upper)) +
  median_hline + 
  geom_crossbar(alpha=0.5, color="white") +
  labs(title="Age", x="Age Range", y="Salary Range") + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  guides(fill="none") +
  coord_flip()
```

```{r experience_hist, include=FALSE, echo=FALSE, message=FALSE}
phackers %>% 
  ggplot(aes(x=exp_f, fill=..count..)) + 
  geom_histogram(stat="count") + 
  labs(title="Years of Experience", x="Years of Experience", y="") +
  guides(fill="none")
```

```{r experience_plot, echo=FALSE, message=FALSE}
phackers %>% group_by(exp_f) %>% 
  summarise(salary_median=median(salary_numeric), lower=quantile(salary_numeric, 0.25), upper=quantile(salary_numeric, 0.75)) %>% 
  ggplot(aes(x=exp_f, fill=exp_f, y=salary_median, ymin=lower, ymax=upper)) + 
  median_hline + 
  geom_crossbar(alpha=0.5, color="white")  + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  labs(title="Years of Experience", x="Years of Experience", y="Salary Range") +
  guides(fill="none") + 
  coord_flip()
```


## Education

```{r education_summary, echo=FALSE, message=FALSE}
education_summary <- phackers %>% filter(university != "(My university is not listed)") %>%
  group_by(university) %>% 
  summarise(n=n(), pct=n()/sum(count(phackers)), 
            salary_median=median(salary_numeric), 
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, 0.75))

top_univerities <- education_summary %>% 
  arrange(desc(n)) %>% 
  top_n(10, n)

top_uni_salary <- education_summary %>% arrange(desc(salary_median))
```
### University
Respondents came from many different universities, with `r top_univerities$university[1]` having the most respondents (`r percent(top_univerities$pct[1])`). 

```{r education_plot, echo=FALSE, message=FALSE, include=TRUE}
top_univerities %>% 
  ggplot(aes(x=university, fill=university, y=salary_median, ymin=lower, ymax=upper)) +
  median_hline + 
  geom_crossbar(alpha=0.5, color="white")  + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  labs(title="University Attended", x="", y="Salary Range") +
  guides(fill="none") + 
  coord_flip()
```

### Course Selection and Advanced Degrees

```{r educ_course_degrees, echo=FALSE}
educ_masters <- phackers %>% 
  filter(educ_masters == TRUE) %>% 
  summarise(n=n(), 
            pct=n()/sum(count(phackers)),
            median_salary=median(salary_numeric), 
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, 0.75))
educ_masters$statement <- "I have (completed) a master's degree"

educ_computer_science <- phackers %>% 
  filter(educ_computer_science == TRUE) %>% 
  summarise(n=n(), 
            pct=n()/sum(count(phackers)),
            median_salary=median(salary_numeric), 
            lower=quantile(salary_numeric, 0.25), 
            upper=quantile(salary_numeric, 0.75)) 
educ_computer_science$statement <- "My academic specialty is/was CS"

educ_math_stat_physics <- phackers %>% filter(educ_math_stat_physics == TRUE) %>% 
  summarise(n=n(), 
            pct=n()/sum(count(phackers)),
            median_salary=median(salary_numeric), lower=quantile(salary_numeric, 0.25), upper=quantile(salary_numeric, 0.75)) 
educ_math_stat_physics$statement <- "My academic specialty was math, stats or physics"
educ_it_mis <- phackers %>% filter(educ_it_mis == TRUE) %>% 
  summarise(n=n(), 
            pct=n()/sum(count(phackers)),
            median_salary=median(salary_numeric), lower=quantile(salary_numeric, 0.25), upper=quantile(salary_numeric, 0.75))
educ_it_mis$statement <- "My academic specialty was IT or MIS"
educ_statements <- rbind(educ_masters, educ_computer_science, educ_math_stat_physics, educ_it_mis)
```

The majority (`r percent(mean(phackers$educ_computer_science | phackers$educ_it_mis))`) came from CS, IT or MIS courses. Only `r percent(educ_statements$pct[educ_statements$statement == "I have (completed) a master's degree"])` of respondents had completed a master's degree, and `r percent(educ_statements$pct[educ_statements$statement == "My academic specialty was math, stats or physics"])` came from a mathematics, statistics or physics course.

```{r educ_statements_plot, echo=FALSE}

ggplot(data=educ_statements, aes(x=statement, fill=statement, y=median_salary, ymin=lower, ymax=upper)) +
  median_hline + 
  geom_crossbar(alpha=0.5, color="white")  + 
  scale_y_continuous(breaks = seq(20000,200000,by=20000), labels=function(sal) {sprintf("%.0fK", sal / 1000)}) + 
  labs(title="Course selection and advanced degrees", x="", y="Salary Range") +
  guides(fill="none") + 
  coord_flip()
```
