---
title: "Phackers Salary Survey, part 1"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

# Introduction

The second Philippine Tech Hackers Salary Survey was conducted online using Google Forms. 
A total of 507 community members responded, from companies both large and small, and from a wide variety of
industries. Respondents were mostly software developers, but other professionals who program and are involved in IT also participated. Outliers, and those who reported as living/working outside the country were excluded from the analysis.

The median monthly base salary for the remaining sample was Php 47,500, with the middle half of all respondents earning between Php 27,500 and Php 82,500
```{r echo=FALSE, message=FALSE}
library(doMC)
registerDoMC(cores = 5)

library(tidyverse)
library(scales)  # allows us to use labels=comma in scales

source("salary_survey_util.R")

# read_csv is a tidyverse version, found in the readr library
phackers_raw <- read_csv("Philippine Tech Hackers Salary Survey 2017.csv")
phackers <- preprocess_salary_survey(phackers_raw)
summary(phackers$salary_numeric)
print(paste("Total non-outlier respondents: ", dim(phackers)[1]))
ggplot(phackers, aes(salary_level))  + coord_flip() + geom_bar() + xlab("Total Salary (Philippine Pesos)") + ylab("Respondents") + labs(title="Total Salary")
```

# Location

The overwhelming majority of respondents live and work in Metro Manila, with just over 16% of the respondents from other cities in the Philippines. 
The median salary for respondents living and working in Metro Manila is Php 47,500 with the median outside of NCR Php 10,000 lower at Php 37,500.

```{r, echo=FALSE}
location_summary <- phackers %>% filter(!is.na(ncr)) %>% group_by(ncr) %>% summarise(Median=median(salary_numeric))
phackers %>% filter(!is.na(ncr)) %>% ggplot(aes(x=ncr, y=salary_numeric)) + geom_boxplot() + xlab("Location") + ylab("Salary")
```

# Company Types

The salary survey included questions on industry, company size, and company age. Software was the most well-represented industry (36.44%), followed by IT Consulting (13.78%) and Business Process Outsourcing (6.44%). 

In terms of salary range, Banking/Finance, Advertising/Marketing/PR and Carriers/Telecommunications had the highest median salary ranges at Php 57,000 while
Education had the lowest at Php 32,500 or Php 15,000 below the overall median of Php 47,500.

```{r, echo=FALSE}
top_industry <- phackers %>% filter(employment_f=="Employed") %>% group_by(company_industry_f) %>% 
  summarise(median=median(salary_numeric), n=n(), pct=round(100*n()/sum(count(phackers)),2)) %>% arrange(desc(pct)) %>%
  rename("Industry"=company_industry_f, percentage=pct)
top_ten_industries <- top_industry[1:10,]
# print(top_ten_industries)

# phackers %>% filter(employment_f=="Employed") %>% group_by(company_industry_f) %>%  ggplot(aes(x=company_industry_f, y=salary_numeric)) + geom_boxplot() + coord_flip() + xlab("Industry") + ylab("Salary Range")

# Use inner_join to get the subset phackers in the top 10 industries
inner_join(phackers, top_ten_industries, by=c("company_industry_f" = "Industry")) %>% filter(employment_f=="Employed") %>% 
  ggplot(aes(x=company_industry_f, y=salary_numeric)) + geom_boxplot() + coord_flip() + xlab("Industry") + ylab("Salary Range") + labs(title="Top 10 Industries\nby number of respondents") 
```

Half the respondents work in companies with 100 or less employees, with 22.44% of respondents working in companies with between 2 and 25 employees. Respondents working in very large companies (over 10,000 employees) made up only 7.56% of the sample. 

```{r, echo=FALSE}
company_size_summary <- phackers %>% filter(employment_f == "Employed") %>% group_by(company_size_f) %>% summarise(median(salary_numeric), n=n(), pct=round(100*n()/sum(count(phackers)),2))
phackers %>% filter(employment_f == "Employed") %>% ggplot(aes(y=salary_numeric, x=company_size_f)) + geom_boxplot() + coord_flip() + labs(title="Company Size", x="Company Size", y="Salary Range")
```